GMU MS SWE Handbook\\{\large A Handbook For the Masters of Software Engineering program at George Mason University

### 2024-10-16 ennv.tex
update

@@

\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Tab.}
\makeatletter
\renewcommand{\chapterautorefname}{\S\@gobble}
\renewcommand{\sectionautorefname}{\S\@gobble}
\renewcommand{\subsectionautorefname}{\S\@gobble}
\renewcommand{\appendixautorefname}{\S\@gobble}
\makeatother

@@

\chapter{Search Algorithms}

>>>>

\usepackage[linesnumbered,ruled,vlined]{algorithm2e}

\SetAlgoCaptionSeparator{.}
\SetAlCapFnt{\footnotesize\rm}
\SetAlCapNameFnt{\footnotesize\rm}

\newcommand{\functiontextformat}[1]{\textrm{\texttt{#1}}}

% \newtoggle{usecomment}
% \settoggle{usecomment}{true}

% \newcommand{\tcomment}[1]{\iftoggle{usecomment}{#1}{}}
% \newcommand{\tvn}[1]{\iftoggle{usecomment}{{\color{red}{[TVN]: #1}}}{}}

% \newcommand{\mbd}[1]{\iftoggle{usecomment}{{\color{magenta}{[MBD]: #1}}}{}}
% \newcommand{\matt}[1]{\iftoggle{usecomment}{{\color{magenta}{[MBD]: #1}}}{}}

% \newcommand{\hd}[1]{\iftoggle{usecomment}{{\color{blue}{[HD]: #1}}}{}}

% \newcommand{\mynew}[1]{\iftoggle{usecomment}{{\color{red}{[NEW]: #1}}}{}}

%\newcommand{\ignore}[1]{}

@@

% for captions
\def\Section{\S}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\renewcommand{\algorithmcfname}{Alg.}

% for cross references
% \renewcommand{\algorithmcflinename}{L.}
\renewcommand{\algorithmautorefname}{Alg.}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Tab.}
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\chapterautorefname}{\S\@gobble}
\renewcommand{\sectionautorefname}{\S\@gobble}
\renewcommand{\subsectionautorefname}{\S\@gobble}
\renewcommand{\appendixautorefname}{\S\@gobble}

\newcommand{\nnproofgen}{\texttt{APTPgen}}
\newcommand{\nnproofcheck}{\texttt{APTPcheck}}
\newcommand{\nnproofchecker}{\texttt{APTPchecker}}
\newcommand{\nnproofformat}{\texttt{APTP}}
%\newcommand{\tool}{\texttt{NNProofChecker}}
\newcommand{\veristable}{\texttt{VeriStable}}
\newcommand{\crown}{\texttt{$\alpha\beta$-CROWN}}
\newcommand{\nnenum}{\texttt{nnenum}}
\newcommand{\marabou}{\texttt{Marabou}}
\newcommand{\eran}{\texttt{ERAN}}
\newcommand{\reluplex}{\texttt{Reluplex}}
\newcommand{\reluval}{\texttt{Reluval}}
\newcommand{\neurify}{\texttt{Neurify}}
\newcommand{\nnv}{\texttt{NNV}}
\newcommand{\ovalnnv}{\texttt{OVAL}}
\newcommand{\dnnv}{\texttt{DNNV}}
\newcommand{\verinet}{\texttt{VeriNet}}
\newcommand{\mnbab}{\texttt{MN-BaB}}
% \newcommand{\vnncomp}{VNN-COMP'23}
\newcommand{\planet}{\texttt{Planet}}
\newcommand{\dd}{\texttt{BaB$_{\text{NV}}$}}
\newcommand{\neuralsat}{\texttt{NeuralSAT}}
\newcommand{\mipverify}{\texttt{MIPVerify}}
\newcommand{\crowndefault}{\texttt{$\alpha\beta$-CROWN} (default)}

\makeatletter
\makeatother

@@

\chapter{Search Algorithms}
\autoref{fig:alg} shows \dd{}, a reference architecture~\cite{nakagawa2014consolidating} 
for modern DNN verifiers that we use to illustrate our observations.  
\dd{} takes as input a ReLU-based DNN $\mathcal{N}$ and a formulae $\phi_{in}\Rightarrow \phi_{out}$ representing the property of interest.

\dd{} iterates between two components: \texttt{Decide} (branching, \autoref{line:decide}), which decides (assigns) an activation status value for a neuron, and \texttt{Deduce} (bounding, \autoref{line:deduce}), which checks the feasibility of the current activation pattern. 
To add proof generation capability, \dd{} is instrumented with a proof tree (\texttt{proof}) variable (\autoref{line:prooftree}) to record these branching decisions. The proof is represented as a binary tree structure, where each node represents a neuron and its left and right edges represent its activation decision (active or inactive). %The proof tree is then used to generate a proof in the \nnproofformat{} format (\autoref{sec:proof-format}).

\begin{figure}[t]
    \begin{minipage}[b]{\linewidth}
        \centering
        \begin{minipage}[t]{0.48\textwidth}
            \centering  
            \includegraphics[width=\linewidth]{figure/proof_net.pdf}
            \caption*{(a)}
        \end{minipage}
        \begin{minipage}[t]{0.48\textwidth}
            \centering
            \includegraphics[width=\linewidth]{figure/proof_tree.pdf}
            \caption*{(b)}
        \end{minipage}
        \caption{(a) A simple DNN.  (b) A proof tree produced verifying the property $(x_1, x_2) \in [-2.0, 2.0] \times [-1.0, 1,0] \Rightarrow (y_1 > y_2)$.}
        \label{fig:example}
    \end{minipage}
\end{figure}
%White nodes correspond to branching nodes where \neuralsat{} makes decisions to split ReLU neurons.

\paragraph{Example} \autoref{fig:example}a illustrates a DNN and how \dd{} determines unsatisfiability (i.e., verifies the problem) and generates the unsat proof in \autoref{fig:example}b.

First, \dd{} initializes the activation pattern set \functiontextformat{ActPatterns} with an empty activation pattern $\emptyset$. Then \dd{} enters a loop (\autoref{line:dpllstart}-\autoref{line:dpllend}) to search for a satisfying assignment or a proof of unsatisfiability. In the first iteration, \dd{} selects the only available activation pattern $\emptyset \in \functiontextformat{ActPatterns}$. 
It calls \functiontextformat{Deduce} to check the feasibility of the problem based on the current activation pattern. \functiontextformat{Deduce} uses abstraction to approximate that from the input constraints the output values are feasible for the given network. 
Since \functiontextformat{Deduce} cannot decide infeasibility, \dd{} randomly selects a neuron to split (\functiontextformat{Decide}). Let us assume that it chooses $v_4$ to split, which essentially means the problem is split into two independent subproblems: one with $v_4$ active and the other with $v_4$ inactive.
\dd{} then adds $v_4$ and $\overline{v_4}$ to \functiontextformat{ActPatterns}.

In the second iteration, \dd{} has two subproblems (that can be processed in parallel). For the first subproblem with $v_4$, \functiontextformat{Deduce} cannot decide infeasibility, so it selects $v_2$ to split. It then conjoins $v_4$ with $v_2$ and then with $\overline{v_2}$ and adds both conjuncts to \texttt{ActPatterns}. 
For the second subproblem with $\overline{v_4}$ inactive, \functiontextformat{Deduce} determines that the problem is unsatisfiable and \dd{} saves the node $v_4$ to the proof tree, as node 3, to indicate one unsatisfiable pattern, i.e., whenever the network has $v_4$ being inactive, the problem is unsatisfiable.

In the third iteration, \dd{} has two subproblems for $v_4 \land v_2$ and $v_4 \land \overline{v_2}$. For the first subproblem, \functiontextformat{Deduce} cannot decide infeasibility, so it selects $v_1$ to split. It then conjoins $v_1$ and then $\overline{v_1}$ to the current activation pattern and adds them to \functiontextformat{ActPatterns}. For the second subproblem, \functiontextformat{Deduce} determines that the problem is unsatisfiable and \dd{} saves the node $v_4 \land \overline{v_2}$ to the proof tree, as node 5.

In the fourth iteration, \dd{} has two subproblems for $v_4 \land v_2 \land v_1$ and $v_4 \land v_2 \land \overline{v_1}$. Both subproblems are determined to be unsatisfiable, and \dd{} saves them to the proof tree as nodes 6 and 7, respectively.

Finally, \dd{} has an empty \texttt{ActPatterns}, stops the search, and returns \texttt{unsat} and the proof tree.
